---
title: "KO07CD34 — Full Paper Analysis Script"
author: "Huttenhain Lab"
date: "`r Sys.Date()`"
output: html_document
---

<!--
KO07CD34_Paper_Script.Rmd
-----------------------------------------------------------------------------
One file, start to finish: optionally the raw Spectronaut report itself
(section 1a, off by default -- see RUN_RAW_IMPORT), otherwise the
already-processed MSstats output (section 1b) -> D9.2 exclusion -> PCA ->
Pearson/Spearman correlation -> Figure 3A (temporal cluster heatmap + trend
lines + GO with simplify()) -> volcano plots (raw + adjusted p) -> GO
Up/Down bar plots with simplify() -> Figure 4A (HPA nuclear/lamina heatmap)
-> every output table. Figure3A_Cluster_Heatmap_GO_noD9.2.R,
Figure4A_Nuclear_Heatmap.R, and GO_Simplify_UpDown.R are `source()`d rather
than re-typed, so each stays a single source of truth and prints its own
"Saved: <path>" lines. Everything else (PCA, correlation, volcano, output
tables) is written directly in this file.

Requirements: R 4.x, plus the CRAN/Bioconductor packages listed in this
folder's README.md. Needs network access at knit time to source() the
Huttenhain Lab's shared bp_utils/drb_utils helper scripts from GitHub, and to
query UniProt / org.Hs.eg.db for gene-name mapping.

Usage:
    rmarkdown::render("KO07CD34_Paper_Script.Rmd")   # from this folder
    # or open in RStudio and knit top to bottom.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# All packages, paths, constants, and the shared color palette live here --
# every step below assumes this has already run.
source("R/00_setup.R")
source("R/01_data_loading.R")
source("R/02_pca_functions.R")
source("R/03_correlation_functions.R")
source("R/04_temporal_cluster_functions.R")
source("R/05_go_enrichment_functions.R")
source("R/06_volcano_functions.R")
```

# 1. Load data from the mass spec / MSstats output

One switch controls the source. **`TRUE`**: raw Spectronaut report through
MSstats yourself (section 1a) -- needs network access and the raw `.tsv`
file at `spectronaut.lfq.report` below. **`FALSE`** (default): the four
already-computed MSstats output CSVs in `data_dir` (section 1b).

```{r raw-import-toggle}
RUN_RAW_IMPORT <- FALSE
```

## 1a. (Optional) Import raw Spectronaut data through MSstats

Only runs if `RUN_RAW_IMPORT` is `TRUE`: raw Spectronaut report ->
contaminant removal -> `SpectronauttoMSstatsFormat()` -> `dataProcess()` ->
gene name mapping -> `groupComparison()`.

```{r raw-import-packages, eval = RUN_RAW_IMPORT}
suppressPackageStartupMessages({
  library(devtools)
  library(UniProt.ws)
  library(R.utils)
})
```

```{r raw-import-utils, eval = RUN_RAW_IMPORT}
# Huttenhain Lab shared utility functions, sourced live from GitHub:
# Create.Pipeline.Directories(), Save.Csv.With.Timestamp(),
# Make.LFQ.Contrast.Matrix(), Spectronaut cleaning/enrichment helpers.
utils <- "https://raw.githubusercontent.com/HuttenhainLab/bp_utils/master/"
source(paste0(utils, "enrichmentTestFunctions.R"))
source(paste0(utils, "MSstats_V4_Functions.R"))

drb_utils <- "https://raw.githubusercontent.com/HuttenhainLab/drb_utils/main/"
source(paste0(drb_utils, "data_management.R"))
source(paste0(drb_utils, "spectronaut_data_cleaning.R"))
source(paste0(drb_utils, "gene_ontology_enrichment.R"))

# Bind gene names onto a data table from its UniProt ID column.
map.gene.names <- function(protein.column, taxonomy.id = 9606) {
  unique.phosphosites <- unique(unlist(strsplit(as.character(protein.column), ";")))
  UniProt.IDs <- unique(sapply(unique.phosphosites, function(x) strsplit(x, "_")[[1]][1]))

  UniProt.mapping <- AnnotationDbi::select(UniProt.ws(taxId = 9606),
                       keys = UniProt.IDs,
                       columns = c("UniProtKB", "gene_primary"),
                       keytype = "UniProtKB")

  gene.name.map <- setNames(UniProt.mapping$Gene.Names..primary., UniProt.mapping$From)

  map_protein <- function(entry) {
    phosphosites <- unlist(strsplit(entry, ";"))
    mapped_sites <- sapply(phosphosites, function(site) {
      parts <- unlist(strsplit(site, "_"))
      UniProtID <- parts[1]
      gene_name <- gene.name.map[UniProtID]
      if (!is.na(gene_name)) {
        paste(c(gene_name, parts[-1]), collapse = "_")
      } else {
        site
      }
    })
    paste(mapped_sites, collapse = ";")
  }

  mapped_column <- sapply(as.character(protein.column), map_protein)
  return(mapped_column)
}

map.gene.names.clusterProfiler <- function(protein.column, protein.database.name = "org.Hs.eg.db") {
  unique.phosphosites <- unique(unlist(strsplit(as.character(protein.column), ";")))
  UniProt.IDs <- unique(sapply(unique.phosphosites, function(x) strsplit(x, "_")[[1]][1]))

  if (!requireNamespace(protein.database.name, quietly = TRUE)) {
    stop(paste("Package", protein.database.name, "is not installed. Please install it using BiocManager::install()."))
  }

  protein.database <- get(protein.database.name, envir = asNamespace(protein.database.name))

  mapping <- bitr(unique(UniProt.IDs), fromType = "UNIPROT", toType = "SYMBOL", OrgDb = protein.database)
  gene.name.map <- setNames(mapping$SYMBOL, mapping$UNIPROT)

  map_protein <- function(entry) {
    phosphosites <- unlist(strsplit(entry, ";"))
    mapped_sites <- sapply(phosphosites, function(site) {
      parts <- unlist(strsplit(site, "_"))
      UniProtID <- parts[1]
      gene_name <- gene.name.map[UniProtID]
      if (!is.na(gene_name)) {
        paste(c(gene_name, parts[-1]), collapse = "_")
      } else {
        site
      }
    })
    paste(mapped_sites, collapse = ";")
  }

  mapped_column <- sapply(as.character(protein.column), map_protein)
  return(mapped_column)
}
```

```{r raw-import-parameters, eval = RUN_RAW_IMPORT}
# Data Type: "SP" = Spectronaut, "MQ" = MaxQuant
export.type <- "SP"

# Appended to output folder/file names
data.name <- "KO07_53min"
Create.Pipeline.Directories(data.name)

# TRUE: pipeline makes the replicate column unique by concatenating
# Condition & Replicate, so MSstats doesn't treat technical replicates as
# bioreplicates (correct for this APEX experiment's design).
is.case.control <- TRUE

if (export.type == "SP") {
  # EDIT ME: path to your raw Spectronaut .tsv export.
  spectronaut.lfq.report <- "path/to/KO07_53min_Report.tsv"
} else if (export.type == "MQ") {
  maxquant.evidence      <- "../example_data/example_data_MQ_evidence.txt"
  maxquant.proteingroups <- "../example_data/example_data_MQ_proteinGroups.txt"
  maxquant.annotations   <- "../example_data/example_data_MQ_annotations.csv"
} else {
  stop("Non-supported data type. Make sure data type is SP or MQ, otherwise pipeline will break.")
}

# Every pairwise comparison MSstats will test (a superset of
# comparisons_of_interest in R/00_setup.R -- everything downstream filters
# to whichever subset it needs).
global.regexContrasts <- c(
  "D14-D5", "D14-D9", "Primary-D14", "Primary-dHL60", "D14-dHL60",
  "Primary-HL60", "Primary-D5", "D9-D5", "Primary-D9", "D14-HL60"
)

# Runs to drop before normalization (e.g. failed injections) -- none by
# default, same as the original script.
runs.to.remove <- c()
```

```{r raw-import-load, eval = RUN_RAW_IMPORT}
if (export.type == "SP") {
  spectronaut.lfq.report <- fread(spectronaut.lfq.report)
} else if (export.type == "MQ") {
  maxquant.evidence      <- fread(maxquant.evidence)
  maxquant.proteingroups <- fread(maxquant.proteingroups)
  maxquant.annotations   <- fread(maxquant.annotations)
}
```

```{r raw-import-clean-contaminants, eval = RUN_RAW_IMPORT}
if (export.type == "SP") {
  # Contaminant protein accessions are prepended with "Cont_" -- Spectronaut
  # doesn't strip these for you.
  spectronaut.lfq.report <- spectronaut.lfq.report %>%
    filter(!grepl("Cont_|NaN|TRYP_PIG", PG.ProteinGroups))
}
```

```{r raw-import-unique-replicates, eval = RUN_RAW_IMPORT}
if (is.case.control) {
  spectronaut.lfq.report$R.Replicate <- paste(spectronaut.lfq.report$R.Condition, spectronaut.lfq.report$R.Replicate, sep = ".")
}
```

```{r raw-import-to-msstats-format, eval = RUN_RAW_IMPORT, warning=FALSE}
if (export.type == "SP") {
  global.prepared <- SpectronauttoMSstatsFormat(
    spectronaut.lfq.report,
    filter_with_Qvalue = TRUE,
    qvalue_cutoff = 0.01,
    removeProtein_with1Feature = FALSE,
    use_log_file = FALSE
  )
  Save.Csv.With.Timestamp(global.prepared, "CleanedPreprocessedData.csv", paste(data.name, "data", sep = "_"))
} else if (export.type == "MQ") {
  global.prepared <- MaxQtoMSstatsFormat(maxquant.evidence, maxquant.annotations, maxquant.proteingroups)
  rm(maxquant.evidence, maxquant.annotations, maxquant.proteingroups)
}

global.prepared <- global.prepared[!(global.prepared$Run %in% runs.to.remove), ]
```

```{r raw-import-dataprocess, eval = RUN_RAW_IMPORT, warning=FALSE}
global.proteinSummarization <- MSstats::dataProcess(
  global.prepared,
  normalization = 'equalizeMedians',
  logTrans = 2,
  featureSubset = 'highQuality',
  summaryMethod = "TMP",
  censoredInt = 'NA',
  MBimpute = FALSE,
  maxQuantileforCensored = 0.999
)

global.proteinSummarization$ProteinLevelData$gene.name <- map.gene.names(global.proteinSummarization$ProteinLevelData$Protein)

Save.Csv.With.Timestamp(global.proteinSummarization$FeatureLevelData, "FeatureLevelData.csv", paste(data.name, "data", sep = "_"))
Save.Csv.With.Timestamp(global.proteinSummarization$ProteinLevelData, "ProteinLevelData.csv", paste(data.name, "data", sep = "_"))
```

```{r raw-import-groupcomparison, eval = RUN_RAW_IMPORT}
contrast.matrix <- Make.LFQ.Contrast.Matrix(input.data.frame = global.proteinSummarization, condition.vector = global.regexContrasts)

global.pairwiseComparison <- MSstats::groupComparison(contrast.matrix, global.proteinSummarization)$ComparisonResult
global.pairwiseComparison$gene.name <- map.gene.names.clusterProfiler(global.pairwiseComparison$Protein)

Save.Csv.With.Timestamp(global.pairwiseComparison, "GroupComparisonsData.csv", paste(data.name, "data", sep = "_"))
```

```{r raw-import-handoff, eval = RUN_RAW_IMPORT}
# Feed straight into the same object names the rest of this pipeline (and
# every other script in this project) expects -- skips re-reading whatever
# timestamped filename Save.Csv.With.Timestamp() just wrote, since that
# stamp depends on today's date and isn't something this file can predict
# ahead of time.
protein   <- standardize_protein_dt(as.data.table(global.proteinSummarization$ProteinLevelData))
groupcomp <- standardize_groupcomp_dt(as.data.table(global.pairwiseComparison))
feature   <- as.data.table(global.proteinSummarization$FeatureLevelData)
cleaned   <- as.data.table(global.prepared)

cat("Raw import complete -- protein:", nrow(protein), "rows | groupcomp:", nrow(groupcomp),
    "rows | feature:", nrow(feature), "rows | cleaned:", nrow(cleaned), "rows\n")
```

## 1b. Load the already-computed MSstats output (default)

Runs only if `RUN_RAW_IMPORT` is `FALSE` above -- reads the four
pre-computed CSVs from `data_dir` (`R/00_setup.R`), exactly as every other
script in this project does.

```{r load-data, eval = !RUN_RAW_IMPORT}
raw <- load_proteomics_data(data_dir)

protein   <- standardize_protein_dt(raw$protein)     # ProteinLevelData
groupcomp <- standardize_groupcomp_dt(raw$groupcomp)  # GroupComparisonsData
feature   <- raw$feature                              # FeatureLevelData
cleaned   <- raw$cleaned                               # CleanedPreprocessedData
```

```{r load-data-summary}
cat("Protein-level rows:", nrow(protein), " | columns:", paste(names(protein), collapse = ", "), "\n")
cat("Group comparison rows:", nrow(groupcomp), " | columns:", paste(names(groupcomp), collapse = ", "), "\n")
cat("Feature-level rows:", nrow(feature), " | columns:", paste(names(feature), collapse = ", "), "\n")
cat("Cleaned/preprocessed rows:", nrow(cleaned), " | columns:", paste(names(cleaned), collapse = ", "), "\n")
```

# 2. Exclude the D9.2 outlier

`outlier_samples` (currently `r outlier_samples`, set once in `R/00_setup.R`)
is excluded from analysis everywhere below — a single switch instead of
filtering by hand in each section. **The raw data for D9.2 is never deleted
anywhere on disk** — it's still present, untouched, in every file loaded
above; this only filters it out at the analysis step, consistently, so no
figure or table includes it while another quietly still does.

Suggested Methods-section language for this exclusion (see the project's
`README.md` for the fuller version): *"One biological replicate (D9.2) was
excluded from downstream analysis due to a documented contamination event
during differentiation, occurring independently of sample collection or MS
processing; raw data for this replicate were retained but excluded from all
reported analyses."*

```{r qc-filtered-protein-table}
protein_qc <- protein[!(SUBJECT %in% outlier_samples)]
cat("Dropped as outliers:", paste(outlier_samples, collapse = ", "), "\n")
cat("Remaining D9 subjects:", paste(sort(unique(protein_qc[GROUP == "D9", SUBJECT])), collapse = ", "), "\n")
```

# 3. PCA — all samples, then with D9.2 removed

```{r pca-all-samples, fig.width=7, fig.height=6}
intensity_matrix <- build_intensity_matrix(protein)

pca_all <- run_pca(intensity_matrix)
p_pca_all <- plot_pca(pca_all)
p_pca_all

ggsave(file.path(fig_dir, "PCA", "PCA_all_samples.pdf"), p_pca_all, width = 7, height = 6)
cat("Saved:", file.path(fig_dir, "PCA", "PCA_all_samples.pdf"), "\n")
```

```{r pca-outlier-removed, fig.width=7, fig.height=6}
run_removed_matrix <- drop_samples(intensity_matrix, outlier_samples)

pca_removed <- run_pca(run_removed_matrix)
p_pca_removed <- plot_pca(pca_removed)
p_pca_removed

ggsave(file.path(fig_dir, "PCA", "PCA_outlier_removed.pdf"), p_pca_removed, width = 7, height = 6)
cat("Saved:", file.path(fig_dir, "PCA", "PCA_outlier_removed.pdf"), "\n")
```

# 4. Pearson correlation

## 4a. Sample-level Pearson correlation heatmap

```{r pearson-sample-correlation, fig.width=8, fig.height=7}
cor_matrix_samples_pearson <- compute_sample_correlation(run_removed_matrix, method = "pearson")
ht_sample_cor_pearson <- plot_sample_correlation_heatmap(cor_matrix_samples_pearson, name = "Pearson r")
draw(ht_sample_cor_pearson)

save_heatmap(ht_sample_cor_pearson,
             file.path(fig_dir, "Correlation_Heatmaps", "Sample_correlation_heatmap_Pearson.pdf"),
             width = 8, height = 7)
cat("Saved:", file.path(fig_dir, "Correlation_Heatmaps", "Sample_correlation_heatmap_Pearson.pdf"), "\n")
```

## 4b. Pairwise log2FC heatmap (D5/D9/D14 comparisons)

```{r pearson-log2fc-heatmap, fig.width=7, fig.height=8}
keep_labels <- c("D9-D5", "D14-D9", "D14-D5")

heat_df <- unique(
  groupcomp[Label %in% keep_labels & !is.na(gene.name) & !is.na(log2FC),
            .(gene.name, Label, log2FC)],
  by = c("gene.name", "Label")
)
heat_wide <- dcast(heat_df, gene.name ~ Label, value.var = "log2FC", fun.aggregate = mean)
heat_mat_log2fc <- as.matrix(as.data.frame(heat_wide)[, -1])
rownames(heat_mat_log2fc) <- heat_wide$gene.name
heat_mat_log2fc <- heat_mat_log2fc[apply(heat_mat_log2fc, 1, function(x) all(is.finite(x))), , drop = FALSE]

ht_log2fc <- Heatmap(
  heat_mat_log2fc,
  name = "log2FC",
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = FALSE,
  show_column_names = TRUE,
  column_title = "Pairwise changes across D5, D9, D14",
  row_title = paste(nrow(heat_mat_log2fc), "proteins"),
  col = pink_ramp_diverging()
)
draw(ht_log2fc)

save_heatmap(ht_log2fc, file.path(fig_dir, "Correlation_Heatmaps", "Pairwise_log2FC_heatmap.pdf"),
             width = 7, height = 8)
cat("Saved:", file.path(fig_dir, "Correlation_Heatmaps", "Pairwise_log2FC_heatmap.pdf"), "\n")
```

## 4c. Cross-condition Pearson panels (G/H/I style figure)

```{r pearson-ghi-panels, fig.width=14, fig.height=4.8}
scatter_d14   <- compute_condition_scatter_data(protein_qc, "Primary", "D14")
scatter_dhl60 <- compute_condition_scatter_data(protein_qc, "Primary", "dHL60")

all_vals <- c(scatter_d14$x, scatter_d14$y, scatter_dhl60$x, scatter_dhl60$y)
axis_lim <- c(floor(min(all_vals, na.rm = TRUE)), ceiling(max(all_vals, na.rm = TRUE)))

p1 <- plot_condition_scatter(scatter_d14,  "Primary", "D14",   axis_lim, axis_lim, line_color = "#2C8ED6")
p2 <- plot_condition_scatter(scatter_dhl60, "Primary", "dHL60", axis_lim, axis_lim, line_color = "#F4A022")

cond_keep <- c("D5", "HL60", "dHL60", "Primary", "D14", "D9")
cor_mat_cond_pearson <- compute_condition_correlation(protein_qc, cond_keep, method = "pearson")

manual_order <- cond_keep[cond_keep %in% colnames(cor_mat_cond_pearson)]
p3_manual <- plot_condition_correlation_tile(cor_mat_cond_pearson, order_type = "manual", order = manual_order)
p3_auto   <- plot_condition_correlation_tile(cor_mat_cond_pearson, order_type = "cluster")

p1_lab  <- add_panel_label(p1, "G)")
p2_lab  <- add_panel_label(p2, "H)")
p3m_lab <- add_panel_label(p3_manual, "I)")
p3a_lab <- add_panel_label(p3_auto, "I)")

combined_manual <- p1_lab + p2_lab + p3m_lab + plot_layout(widths = c(1, 1, 1.05))
combined_auto   <- p1_lab + p2_lab + p3a_lab + plot_layout(widths = c(1, 1, 1.05))
print(combined_manual)
print(combined_auto)

ggsave(file.path(fig_dir, "Correlation_Heatmaps", "Pearson_GHI_manual_order.pdf"), combined_manual, width = 14, height = 4.8)
ggsave(file.path(fig_dir, "Correlation_Heatmaps", "Pearson_GHI_auto_order.pdf"), combined_auto, width = 14, height = 4.8)
ggsave(file.path(fig_dir, "Correlation_Heatmaps", "Pearson_GHI_manual_order.png"), combined_manual, width = 14, height = 4.8, dpi = 300)
ggsave(file.path(fig_dir, "Correlation_Heatmaps", "Pearson_GHI_auto_order.png"), combined_auto, width = 14, height = 4.8, dpi = 300)
cat("Saved GHI panels to:", file.path(fig_dir, "Correlation_Heatmaps"), "\n")
```

## 4d. Condition correlation heatmap with dendrograms

```{r pearson-condition-dendro, fig.width=6.5, fig.height=5.5}
condition_colors <- c(
  "D5" = "#8DB6D9", "HL60" = "#F2AE4A", "dHL60" = "#E49A2F",
  "Primary" = "#C03A87", "D14" = "#2C8ED6", "D9" = "#4F93C6"
)

ht_condition_dendro_pearson <- plot_condition_correlation_dendro(cor_mat_cond_pearson, condition_colors)
draw(ht_condition_dendro_pearson, heatmap_legend_side = "right", annotation_legend_side = "right")

save_heatmap(ht_condition_dendro_pearson,
             file.path(fig_dir, "Correlation_Heatmaps", "Pearson_condition_matrix_dendro.pdf"),
             width = 6.5, height = 5.5,
             heatmap_legend_side = "right", annotation_legend_side = "right")
cat("Saved:", file.path(fig_dir, "Correlation_Heatmaps", "Pearson_condition_matrix_dendro.pdf"), "\n")
```

# 5. Spearman correlation

Same two panels as Pearson above (sample-level + condition-level), as a
robustness check with rank-based correlation instead of linear.

## 5a. Sample-level Spearman correlation heatmap

```{r spearman-sample-correlation, fig.width=8, fig.height=7}
cor_matrix_samples_spearman <- compute_sample_correlation(run_removed_matrix, method = "spearman")
ht_sample_cor_spearman <- plot_sample_correlation_heatmap(cor_matrix_samples_spearman, name = "Spearman rho")
draw(ht_sample_cor_spearman)

save_heatmap(ht_sample_cor_spearman,
             file.path(fig_dir, "Correlation_Heatmaps", "Sample_correlation_heatmap_Spearman.pdf"),
             width = 8, height = 7)
cat("Saved:", file.path(fig_dir, "Correlation_Heatmaps", "Sample_correlation_heatmap_Spearman.pdf"), "\n")
```

## 5b. Condition-level Spearman correlation (tile heatmap)

```{r spearman-condition-correlation, fig.width=7, fig.height=6}
cor_mat_cond_spearman <- compute_condition_correlation(protein_qc, cond_keep, method = "spearman")

p_spearman_manual <- plot_condition_correlation_tile(cor_mat_cond_spearman, order_type = "manual", order = manual_order)
p_spearman_auto   <- plot_condition_correlation_tile(cor_mat_cond_spearman, order_type = "cluster")
print(p_spearman_manual)
print(p_spearman_auto)

ggsave(file.path(fig_dir, "Correlation_Heatmaps", "Spearman_condition_manual_order.pdf"), p_spearman_manual, width = 7, height = 6)
ggsave(file.path(fig_dir, "Correlation_Heatmaps", "Spearman_condition_auto_order.pdf"), p_spearman_auto, width = 7, height = 6)
ggsave(file.path(fig_dir, "Correlation_Heatmaps", "Spearman_condition_manual_order.png"), p_spearman_manual, width = 7, height = 6, dpi = 300)
ggsave(file.path(fig_dir, "Correlation_Heatmaps", "Spearman_condition_auto_order.png"), p_spearman_auto, width = 7, height = 6, dpi = 300)
cat("Saved Spearman condition-level heatmaps to:", file.path(fig_dir, "Correlation_Heatmaps"), "\n")
```

# 6. Figure 3A — temporal cluster heatmap, trend lines, GO (simplify()-collapsed)

D9.2 excluded (`k = 4`, this script's own local override — see its header
comment for why). Runs the standalone script's full pipeline top to bottom
and saves its own heatmap PDF/PNG, translation tables (heatmap-cluster-
number-to-table-cluster-number), GO term tables, and per-protein z-score
matrix — watch the "Saved:" lines in the output below for every file this
step produces.

```{r figure-3a-d92-excluded}
source("Figure3A_Cluster_Heatmap_GO_noD9.2.R")
```

# 7. Volcano plots

## 7a. Raw p-value axis

```{r volcano-auto-labeled, fig.width=8, fig.height=10}
volc_x_lim <- c(-4.5, 8.5)
volc_y_lim <- c(0, 12.5)
comparisons_updown <- c("D9-D5", "D14-D9")

p_volc_B_auto <- plot_volcano("D9-D5", groupcomp, label_mode = "top_n", top_n = 20,
                               title = "D9 vs D5")
p_volc_C_auto <- plot_volcano("D14-D9", groupcomp, label_mode = "top_n", top_n = 20,
                               title = "D14 vs D9")
combined_volcano_auto <- p_volc_B_auto / p_volc_C_auto
print(combined_volcano_auto)

ggsave(file.path(fig_dir, "Volcano_Plots", "D9D5_D14D9_volcano_autolabeled.pdf"),
       combined_volcano_auto, width = 8, height = 10)
```

```{r volcano-custom-labeled, fig.width=8, fig.height=10}
genes_B <- c("DEFA1", "FCAR", "ITGAM", "CYBB", "PADI4",       # D9-D5 up
             "TAL1", "HBE1", "HLA-DRA", "SYTL4", "SNCA")      # D9-D5 down
genes_C <- c("DAB2", "ANXA2", "SDCBP", "ICAM1", "PDLIM2",     # D14-D9 up
             "NOL6", "RPS4Y1", "NDUFA3", "TIMM23", "ESRRA")   # D14-D9 down

p_volc_B <- plot_volcano("D9-D5", groupcomp, label_mode = "custom", label_genes = genes_B,
                          x_lim = volc_x_lim, y_lim = volc_y_lim,
                          title = "Early Neutrophils vs Granulocytes",
                          x_label = "Log2FC (Early Neutrophils/Granulocytes)")
p_volc_C <- plot_volcano("D14-D9", groupcomp, label_mode = "custom", label_genes = genes_C,
                          x_lim = volc_x_lim, y_lim = volc_y_lim,
                          title = "Mature Neutrophils vs Early Neutrophils",
                          x_label = "Log2FC (Mature Neutrophils/Early Neutrophils)")
combined_volcano_custom <- p_volc_B / p_volc_C
print(combined_volcano_custom)

ggsave(file.path(fig_dir, "Volcano_Plots", "D9D5_D14D9_volcano_labeled.pdf"),
       combined_volcano_custom, width = 8, height = 10)
ggsave(file.path(fig_dir, "Volcano_Plots", "D9D5_D14D9_volcano_labeled.png"),
       combined_volcano_custom, width = 8, height = 10, dpi = 300)
ggsave(file.path(fig_dir, "Volcano_Plots", "D9D5_volcano_labeled.pdf"), p_volc_B, width = 8, height = 5)
ggsave(file.path(fig_dir, "Volcano_Plots", "D14D9_volcano_labeled.pdf"), p_volc_C, width = 8, height = 5)
```

```{r volcano-d14d5-plus-go, fig.width=12, fig.height=12}
p_volc_D14D5 <- plot_volcano(
  "D14-D5", groupcomp,
  label_mode = "top_n", top_n = 20,
  x_lim = volc_x_lim, y_lim = volc_y_lim,
  title = "D14 vs D5"
)

gene_universe_D14D5 <- unique(groupcomp[Label == "D14-D5" & !is.na(gene.name), gene.name])
go_D14D5 <- plot_go_updown(
  "D14-D5", groupcomp, gene_universe_D14D5,
  padj_cutoff = PADJ_CUTOFF, top_n = 10
)

if (!is.null(go_D14D5$plot)) {
  combined_D14D5_volcano_go <- p_volc_D14D5 + go_D14D5$plot + plot_layout(widths = c(1, 1.3))
} else {
  combined_D14D5_volcano_go <- p_volc_D14D5
  message("No GO BP terms passed the cutoff for D14-D5 up/down genes; showing volcano only.")
}
print(combined_D14D5_volcano_go)

ggsave(file.path(fig_dir, "Volcano_Plots", "D14D5_volcano_with_top10_GO.pdf"),
       combined_D14D5_volcano_go, width = 12, height = 12)
ggsave(file.path(fig_dir, "Volcano_Plots", "D14D5_volcano_with_top10_GO.png"),
       combined_D14D5_volcano_go, width = 12, height = 12, dpi = 300)
ggsave(file.path(fig_dir, "Volcano_Plots", "D14D5_volcano_labeled.pdf"),
       p_volc_D14D5, width = 8, height = 5)
ggsave(file.path(fig_dir, "Volcano_Plots", "D14D5_volcano_labeled.png"),
       p_volc_D14D5, width = 8, height = 5, dpi = 300)
```

Top hits tables (one Up/Down table per comparison):

```{r volcano-top-hits-tables}
for (comp in comparisons_updown) {
  hits <- top_hits_table(groupcomp, comp, padj_cutoff = PADJ_CUTOFF, n = 50)
  cat("\n====", comp, "====\nUP:\n"); print(hits$up)
  cat("\nDOWN:\n"); print(hits$down)
}
```

## 7b. Adjusted p-value axis

Same plots, re-drawn with `-log10(adjusted p-value)` on the y-axis instead of
raw p-value; axis bounds are computed fresh since adjusted p-values are
always >= raw (BH correction only inflates them).

```{r volcano-shared-limits-adjp}
compute_shared_volcano_limits <- function(groupcomp_dt, labels, y_metric = "adj.pvalue") {
  sub <- groupcomp_dt[
    Label %in% labels & !is.na(log2FC) & !is.na(pvalue) & !is.na(adj.pvalue)
  ]
  yvals <- -log10(sub[[y_metric]])
  list(
    x_lim = c(floor(min(sub$log2FC, na.rm = TRUE)), ceiling(max(sub$log2FC, na.rm = TRUE))),
    y_lim = c(0, ceiling(max(yvals, na.rm = TRUE)))
  )
}

out_dir_adjp <- file.path(fig_dir, "Adjusted_Pvalue_Test_Figures")
dir.create(out_dir_adjp, recursive = TRUE, showWarnings = FALSE)

adj_limits <- compute_shared_volcano_limits(groupcomp, c("D9-D5", "D14-D9", "D14-D5"))
adj_x_lim <- adj_limits$x_lim
adj_y_lim <- adj_limits$y_lim
cat("Shared adjusted-p-value axis bounds -- x:", adj_x_lim, " y:", adj_y_lim, "\n")
```

```{r volcano-auto-labeled-adjp, fig.width=8, fig.height=10}
p_volc_B_auto_adjp <- plot_volcano(
  "D9-D5", groupcomp,
  label_mode = "top_n", top_n = 20,
  x_lim = adj_x_lim, y_lim = adj_y_lim,
  y_metric = "adj.pvalue",
  title = "D9 vs D5"
)
p_volc_C_auto_adjp <- plot_volcano(
  "D14-D9", groupcomp,
  label_mode = "top_n", top_n = 20,
  x_lim = adj_x_lim, y_lim = adj_y_lim,
  y_metric = "adj.pvalue",
  title = "D14 vs D9"
)
combined_volcano_auto_adjp <- p_volc_B_auto_adjp / p_volc_C_auto_adjp
print(combined_volcano_auto_adjp)

ggsave(file.path(out_dir_adjp, "D9D5_D14D9_volcano_autolabeled_ADJP.pdf"),
       combined_volcano_auto_adjp, width = 8, height = 10)
```

```{r volcano-custom-labeled-adjp, fig.width=8, fig.height=10}
p_volc_B_adjp <- plot_volcano(
  "D9-D5", groupcomp, label_mode = "custom", label_genes = genes_B,
  x_lim = adj_x_lim, y_lim = adj_y_lim, y_metric = "adj.pvalue",
  title = "Early Neutrophils vs Granulocytes",
  x_label = "Log2FC (Early Neutrophils/Granulocytes)"
)
p_volc_C_adjp <- plot_volcano(
  "D14-D9", groupcomp, label_mode = "custom", label_genes = genes_C,
  x_lim = adj_x_lim, y_lim = adj_y_lim, y_metric = "adj.pvalue",
  title = "Mature Neutrophils vs Early Neutrophils",
  x_label = "Log2FC (Mature Neutrophils/Early Neutrophils)"
)
combined_volcano_custom_adjp <- p_volc_B_adjp / p_volc_C_adjp
print(combined_volcano_custom_adjp)

ggsave(file.path(out_dir_adjp, "D9D5_D14D9_volcano_labeled_ADJP.pdf"),
       combined_volcano_custom_adjp, width = 8, height = 10)
ggsave(file.path(out_dir_adjp, "D9D5_D14D9_volcano_labeled_ADJP.png"),
       combined_volcano_custom_adjp, width = 8, height = 10, dpi = 300)
ggsave(file.path(out_dir_adjp, "D9D5_volcano_labeled_ADJP.pdf"), p_volc_B_adjp, width = 8, height = 5)
ggsave(file.path(out_dir_adjp, "D14D9_volcano_labeled_ADJP.pdf"), p_volc_C_adjp, width = 8, height = 5)
```

```{r volcano-d14d5-adjp, fig.width=8, fig.height=5}
p_volc_D14D5_adjp <- plot_volcano(
  "D14-D5", groupcomp,
  label_mode = "top_n", top_n = 20,
  x_lim = adj_x_lim, y_lim = adj_y_lim,
  y_metric = "adj.pvalue",
  title = "D14 vs D5"
)
print(p_volc_D14D5_adjp)

ggsave(file.path(out_dir_adjp, "D14D5_volcano_labeled_ADJP.pdf"), p_volc_D14D5_adjp, width = 8, height = 5)
ggsave(file.path(out_dir_adjp, "D14D5_volcano_labeled_ADJP.png"), p_volc_D14D5_adjp, width = 8, height = 5, dpi = 300)
```

```{r volcano-d14d5-plus-go-adjp, fig.width=12, fig.height=7.5}
go_D14D5_adjp <- plot_go_updown(
  "D14-D5", groupcomp, gene_universe_D14D5,
  padj_cutoff = PADJ_CUTOFF, top_n = 5
)

if (!is.null(go_D14D5_adjp$plot)) {
  combined_D14D5_volcano_go_adjp <- p_volc_D14D5_adjp + go_D14D5_adjp$plot + plot_layout(widths = c(1, 1.3))
} else {
  combined_D14D5_volcano_go_adjp <- p_volc_D14D5_adjp
  message("No GO BP terms passed the cutoff for D14-D5 up/down genes; showing volcano only.")
}
print(combined_D14D5_volcano_go_adjp)

ggsave(file.path(out_dir_adjp, "D14D5_volcano_with_top5_GO_ADJP.pdf"),
       combined_D14D5_volcano_go_adjp, width = 12, height = 7.5)
ggsave(file.path(out_dir_adjp, "D14D5_volcano_with_top5_GO_ADJP.png"),
       combined_D14D5_volcano_go_adjp, width = 12, height = 7.5, dpi = 300)
cat("Saved adjusted-p-value volcano plots to:", out_dir_adjp, "\n")
```

# 8. GO Biological Process, Up/Down, simplify()-collapsed

Runs `GO_Simplify_UpDown.R`: per comparison, Up/Down gene sets (`adj.pvalue
< 0.05`, no fold-change cutoff) -> `enrichGO()` -> `simplify()` (Wang, cutoff
0.7) -> top-5-per-direction bar plot. Saves to its own `output/` subfolder
at the project root, not `figures/GO_Enrichment/`.

```{r go-simplify-updown}
source("GO_Simplify_UpDown.R")
```

# 9. Figure 4A — HPA nuclear/lamina heatmap

Runs `Figure4A_Nuclear_Heatmap.R`: HPA nuclear gene list restricted to
significant genes, LBR/LMNA/LMNB1/LMNB2 always forced in and labeled,
`k = 4` clusters, dendrogram heatmap, per-gene trajectory panels, and the
translation-table + per-protein z-score CSV export.

```{r figure-4a-nuclear}
source("Figure4A_Nuclear_Heatmap.R")
```

# 10. Output tables: group comparisons + cleaned/processed data

Saves the four core MSstats tables as a standalone, versioned snapshot, so a
reviewer has the exact tables this run was built from without `data_dir`.
Everything else (cluster gene lists, GO tables, z-score matrices) is already
saved by the sections above.

```{r output-tables-processed-data}
processed_dir <- file.path(table_dir, "processed_data")
dir.create(processed_dir, recursive = TRUE, showWarnings = FALSE)

fwrite(groupcomp, file.path(processed_dir, "GroupComparisons_Table.csv"))

# Full table (D9.2 included) and the QC-filtered version actually used
# throughout the rest of the pipeline, saved separately.
fwrite(protein,    file.path(processed_dir, "ProteinLevelData_all_samples.csv"))
fwrite(protein_qc, file.path(processed_dir, "ProteinLevelData_QC_filtered_noD9.2.csv"))

fwrite(feature, file.path(processed_dir, "FeatureLevelData.csv"))
fwrite(cleaned, file.path(processed_dir, "CleanedPreprocessedData.csv"))

cat("Saved processed data tables to:\n ", processed_dir, "\n")
cat(" -", "GroupComparisons_Table.csv", "(", nrow(groupcomp), "rows )\n")
cat(" -", "ProteinLevelData_all_samples.csv", "(", nrow(protein), "rows )\n")
cat(" -", "ProteinLevelData_QC_filtered_noD9.2.csv", "(", nrow(protein_qc), "rows )\n")
cat(" -", "FeatureLevelData.csv", "(", nrow(feature), "rows )\n")
cat(" -", "CleanedPreprocessedData.csv", "(", nrow(cleaned), "rows )\n")
```

# Session info

```{r session-info}
sessionInfo()
```
